---
title: "Project 2.2"
author: "Oluwakemi Omotunde"
date: "October 09, 2016"
output: html_document
---
``` {r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)

```

##**Task**  

####*We are to choose any three of the "wide" datasets identified in the Week 5 Discussion items and for each:*

+ *Create a .CSV file that includes all of the information included in the dataset. We're encouraged to use a "wide" stucture similar to how th information appears in the discussion item, so we can practice tidying and transformations.*
+ *Read the information from your .CSV file into R, and use tidyr and dplyr as needed to tidy and transform your data.*
+ *Perform the analysis requested in the discussion item.*
+ *Code should be in an R Markdown file, posted to rpubs.com, and should include narrative dscriptions of your dataa clean up work, analysis and conclusions.*

##**Introduction**  

#####*For my second dataset, I decided to take a look at hospitals in the tri-state area(NY, NJ, CT). I would like to be able to determine which state has the best hospitals overall as well as which hosptals have the top ratings.*

##**Hypothesis**

####*


###**Loading data and required libraries**

```{r load_data_libraries}
library(plyr)
library(dplyr)
library(ggplot2)
library(plotly)

hospital <- read.csv("https://raw.githubusercontent.com/komotunde/DATA607/master/Project%202/Hospital%20Info.csv", colClasses = "character")
View(hospital)
```

#####*Since we are only concerned about the hospitals located  in the Tri-State area, our first task is to create a subset that will only include those states. Our main focus is going to be exploring how well a hospital is rated so we will only look at their "achievement points". I would also like to see which category has the highest correlation to HCAHPS scores.*

```{r subset}
hospital <- as.data.frame(hospital[c(2, 5, 8, 11, 14, 17, 20, 23, 26, 29, 32, 33)])  
##this will give me the columns pertaining to acheivement points and HCAHPS base and consistency score

is.data.frame(hospital)  
##i did this just to verify that I have a data frame

colnames(hospital) <- c("Hospital.Name" = "NAME", "State" = "STATE", "Communication.with.Nurses.Achievement.Points" = "COMM.NURSES", "Communication.with.Doctors.Achievement.Points" = "COMM.DOCTORS", "Responsiveness.of.Hospital.Staff.Achievement.Points" = "RESPONSIVENESS", "Pain.Management.Achievement.Points" = "PAIN.MAN", "Communication.about.Medicines.Achievement.Points" = "COMM.MEDS", "Cleanliness.and.Quietness.of.Hospital.Environment.Achievement.Points" = "ENVIRONMENT", "Discharge.Information.Achievement.Points" = "DISCHARGE", "Overall..Rating.of.Hospital.Achievement.Points" = "OVERALL", "HCAHPS.Base.Score" = "HCAHPS.BASE", "HCAHPS.Consistency.Score" = "HCAHPS.CON")

hospital <- subset(hospital, hospital$STATE == 'NY'| hospital$STATE == 'NJ'| hospital$STATE == 'CT')   
##this will result in the data that only pertains to the tri-state area

head(hospital)
```

#####*Before I can do anything mathematically, I need to change the columns from character to numeric. We include the parameter na.rm =TRUE to remove the NA's since we don't want them to affect our calculations.*

```{r ratings numeric}
hospital <- transform(hospital, COMM.NURSES = as.numeric(COMM.NURSES), na.rm = TRUE, warning = FALSE)
hospital <- transform(hospital, COMM.DOCTORS = as.numeric(COMM.DOCTORS), na.rm = TRUE, warning = FALSE)
hospital <- transform(hospital, RESPONSIVENESS = as.numeric(RESPONSIVENESS), na.rm = TRUE, warning = FALSE)
hospital <- transform(hospital, PAIN.MAN = as.numeric(PAIN.MAN), na.rm = TRUE, warning = FALSE)
hospital <- transform(hospital, COMM.MEDS = as.numeric(COMM.MEDS), na.rm = TRUE, warning = FALSE )
hospital <- transform(hospital, ENVIRONMENT = as.numeric(ENVIRONMENT), na.rm = TRUE, warning = FALSE)
hospital <- transform(hospital, DISCHARGE = as.numeric(DISCHARGE), na.rm = TRUE, warning = FALSE)
hospital <- transform(hospital, OVERALL = as.numeric(OVERALL), na.rm = TRUE, warning = FALSE)
hospital <- transform(hospital, HCAHPS.BASE = as.numeric(HCAHPS.BASE), na.rm = TRUE, warning = FALSE)
hospital <- transform(hospital, HCAHPS.CON = as.numeric(HCAHPS.CON), na.rm = TRUE, warning = FALSE)
View(hospital)
```

#####*I'll be using the dplyr and plyr packages to change around our data.*

```{r dplyr/plyr work}
hospital <- mutate(hospital, HCAHPS.TOTAL = HCAHPS.CON + HCAHPS.BASE)
hospital <- hospital[c(-11, -12, -13)]
##I was able to combine both scores for HCAHPS for an overall HCAHPS total and removed the columns not needed. 
hospital <- mutate(hospital, AVE.SCORE = (round(COMM.NURSES + COMM.DOCTORS + RESPONSIVENESS + PAIN.MAN + COMM.MEDS + ENVIRONMENT + DISCHARGE)/7))
hospital <- arrange(hospital, desc(AVE.SCORE))
View(hospital)
##from this we see that the hospital with highest rating is ADIRONDACK MEDICAL CENTER

```

#####*I noticed that there were a few entires where all columns are 0's. I would like to remove those entries.*

```{r remove 0}
hospital <- dplyr::filter(hospital, AVE.SCORE > 0)
View(hospital)
```

#####*I am now at the point whereby my data looks pretty good and I am able to do some analysis and calculations.* 

```{r plot}
hospital.plot <- ggplot(hospital, aes(AVE.SCORE, HCAHPS.TOTAL)) + geom_point() +geom_smooth(method = lm) 
##this regression line includes 95% confidence region
hospital.plot <- hospital.plot + labs(x = "Average Score", y = "HCAHPS SCORE", title = "HCAHPS VS. AVERAGE SCORE")
hospital.plot
##I was expecting that more of the values would be clustered around the line, showing a strong correlation between HCAHPS Scores and Average score overall. This is not the case.
```

#####*Now to see which state has the best ratings overall.*

```{r best state}
hospital.best <- arrange(hospital, STATE, desc(AVE.SCORE), desc(HCAHPS.TOTAL))
##I want to sort my data by state, average score and HCAHPS total score

ct.info <- subset(hospital, STATE == "CT")
ct.average <- mean(ct.info$AVE.SCORE)
ct.HCAHPS <- mean(ct.info$HCAHPS)

ny.info <- subset(hospital, STATE == "NY")
ny.average <- mean(ny.info$AVE.SCORE)
ny.HCAHPS <- mean(ny.info$HCAHPS)

nj.info <- subset(hospital, STATE == "NJ")
nj.average <- mean(nj.info$AVE.SCORE)
nj.HCAHPS <- mean(nj.info$HCAHPS)

average <- data.frame(STATES = c("CT", "NJ", "NY"), AVERAGE = c(1.92, 1.20, 1.56 ))
average
average.plot <- ggplot(average, aes(x = STATES, y = AVERAGE, fill = STATES)) + geom_bar(stat = "identity") + ggtitle("AVERAGE RATING BY STATE")
ggplotly(average.plot)

##We see that Connecticut has the highest average rating overall. We will do the  same thing for HCAPH scores. I would guess that CT will also have the highest HCAPH score.

HCAHPS <- data.frame(STATES = c("CT", "NJ", "NY"), SCORE = c(32.25, 29, 30.3))
HCAHPS
HCAHPS.plot <- ggplot(HCAHPS, aes(x = STATES, y = SCORE, fill = STATES)) + geom_bar(stat = "identity") + ggtitle("HCAHPS SCORE BY STATE")
ggplotly(HCAHPS.plot)

##this plot shows us that CT has the highest HCHPS score of 32.25 out of 100. 

```


##*Conclusion*
We can conclude that Connecticut has the best hospitals in the Tri-State area based on both customer feedback as well as HCAHPS ratings. The best hospital in the Tri-State area is ADIRONDACK MEDICAL CENTER with a HCAHPS score of 89 out of 100 and an average rating of 8.86 out of 10. 

