---
output: pdf_document
---
title: "607 Homework 1"
author: "Oluwakemi Omotunde"
date: "September 10, 2016"
output: pdf_document
---
```{r}
library(RMySQL)
library(plyr)
library(ggplot2)

movierating.upload<- dbConnect(MySQL(), user="root", password="N!cole09", 
    dbname="movies", host="localhost",client.flag=CLIENT_MULTI_STATEMENTS)

dbListTables(movierating.upload)
MovieRatings <- dbReadTable(movierating.upload,"movie_rating")
dbWriteTable(movierating.upload,"movie_rating",MovieRatings,overwrite=T)

MovieRatings
```
#This shows the data frame that was loaded. 

```{r}
MovieRatingsName <- rename(MovieRatings, c("BAD_MOMS"="BadMoms", "SUICIDE_SQUAD"="SuicideSquad", "SUPERMAN_VS_BATMAN"="SupermanVsBatman","CAPTAIN_AMERICA_CIVIL_WAR"="CaptainAmericaCivilWar", "ZOOTOPIA"="Zootopia", "DEADPOOL"="Deadpool"))
```
#This renames some of the columns to make it neater. 

```{r}
MovieRatingsNew <- na.omit(MovieRatingsName)
MovieRatingsNew
```
#I wanted to do this to eliminate the NA fields since I wanted to do some calculations. Upon running this code, I see that it does not serve the purpose I need it to. 

```{r}
mean(MovieRatingsName$BadMoms, na.rm=TRUE) 
```
#This removes the NA fields in calculations and returns a numeric value. 

#To figure out which movie has was favored most, we will plot MovieRatingsNew(no rater who marked NA indicating they haven't seen a particular movie) so that we only consider people who have seen all the movies listed. I believe this will give us more accuarate information.This is working under the assuption that if you've seen all 5 movies, you're a regular movie goer.

```{r}
meanrating <-c(mean(MovieRatingsNew$BadMoms), mean(MovieRatingsNew$SuicideSquad), mean(MovieRatingsNew$SupermanVsBatman), mean(MovieRatingsNew$Zootopia), mean(MovieRatingsNew$Deadpool), mean(MovieRatingsNew$CaptainAmericaCivilWar))

barplot(meanrating)
```
#From here, I would like to make the barplot more meaningful.

```{r}
barplot(meanrating, main= "Average Movie Rating", xlab="Movie", ylab="Average Rating",names.arg=c("BM", "SS", "SvsB","Z", "D", "CACW"))
```
#with the lableling completed, we can see that Deadpool and Captain America have the highest ratings. To check:

```{r}
mean(MovieRatingsNew$Deadpool)
```
#this is shown on the barplot

```{r}
mean(MovieRatingsNew$CaptainAmericaCivilWar)
```
#also shown on the barplot
