---
title: 'Project #1 Re-Do'
author: "Oluwakemi Omotunde"
date: "October 7, 2016"
output: html_document
---

#Introduction  

####After submitting Project 1, I was informed that my submission relied too heavily on Excel due to the fact that I completed the calculations in Excel. Fortunately I was given an opportunity to re-do Project 1 after getting the chance to take a look at other students' completed work.  

#####For this project, we are given a text file with chess tournament results where the information has some structure. We are responsible for creating an R Markdown file that generates a .csv with the following information for all the players:  
  + Player's Name
  + Player's State
  + Total Number of Points
  + Player's Pre-Rating
  + Average Pre-Chess Rating of Opponents  
  
#####We need to calculate the Average Pre Chess Rating of Opponents by adding the pre-rating of each opponent and dividing it by the number of matches played. 

#####Load libraries that are required:
+ 'stringr': we will be using the 'stringr' package for manipulating the data

```{r}
library(stringr)
```

#####Read in the text file:

``` {r read_in_file}
chessinfo <- readLines('chessinfo.txt', warn = FALSE)  
#I get a warning message when this line runs so I added the warn component to eliminate the warning. 
head(chessinfo)
```

#####Once the file has been loaded, we will start manipulating our data by extracting all the pertinent information such as player's names, state and opponent information using str_extract_all().

```{r}
chessinfo <- chessinfo[-c(1:4)]
l <- length(chessinfo)
row.one <- as.factor(chessinfo[seq(1,l,3)])
#we have to create a subset that contains the names

head(row.one)
row.two <- as.factor(chessinfo[seq(2,l,3)])   
#this will contain all the alternate row information, basically just dividing up by every other row

head(row.two)
player.name <- str_trim(str_extract(row.one, "(\\w+\\s){2,3}"))  
#this will extract all words with 2 or more letters, with a space and up to 3 words and then it str_trim will trim down to a new string.

head(player.name)
scores <- as.double(unlist(str_extract_all(row.one, "[:digit:][.][:digit:]")))  
#this will extract all numbers in the format x.x and sets them as doubles 

head(scores)
opponent <- str_extract_all(row.one, "[:digit:]+?\\|") #this is extract all digits followed by |
opponent <- str_extract_all(opponent, "\\d+")
head(opponent)
states <- str_extract(row.two, "\\w+")
head(states)
pre.rating <- str_extract(row.two, "[^[:digit:]][:digit:]{1,4}[^[:digit:]]")
pre.rating <- str_extract(pre.rating, "\\d+")
#I initially tried code that would extract all four digit numbers with spaces before and after but that didn't take into account elements such as 955P11. This approach worked much better

head(pre.rating)
```

#####Now that we have the names, states, scores, and pre-rating extracted, we need to calculated the average opponent pre-rating. We will use a loop to do this. 

```{r}
average.preRating <- 0

for (i in 1:length(player.name)){
  average.preRating[i] <- round(mean(as.numeric(pre.rating[as.numeric(unlist(opponent[i]))]), na.rm = TRUE, digits = 0))
}
head(average.preRating)
```

#####Finally, I will combine all the parts into a data frame, rename the columns and convert to a table.

```{r data frame}
chess.output <- data.frame(player.name, states, scores, pre.rating, average.preRating)
names(chess.output) <- c("Player's Name","State", "Score", "Pre Rating","Ave.Opp.Pre Rating")
chess.output  
knitr::kable(chess.output)
```

#####Now that my data our data is neat and contains our necessary information, we can write a .csv file. 

```{r}
write.csv(chess.output, file = "chess_output.csv")
```

