---
title: "Project 1"
author: "Oluwakemi Omotunde"
date: "September 21, 2016"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##**INSTRUCTIONS**
*In this project, you're given a text file with chess tournament results where the information has some structure. Your job is to create a R Markdown file that generates a .CSV file(that could for example be imported into a SQL database) with the following information for all players:*  

+ Player's Name
+ Player's State
+ Total Number of Points
+ Player's Pre-Rating
+ Average Pre Chess Rating of Opponents  

###**OUTSIDE WORK** 
  
  
*The first thing I did was create an excel spreadsheet that contained each players info along with their scores for each round. I used an excel workbook because I would be able to make updates in the Excel file and once I save, I can run the code again and it will reflect the updates.*

###**R-STUDIO STEPS**

*NoW, we need to set the libraries that we will be utilizing for this project. I went back and added rJava and xlsx.*

```{r}
library(plyr)
library(stringr)
library(xlsx)
library(rJava)
```

*The code below will load our excel file. Next we want to manipulate the file a bit. We will rename the columns first.*

```{r}
df <- read.xlsx("Project 1.xlsx", 1)
chess.info <- df
head(chess.info)
```

*The columns will be renamed for a neater looking file.*


```{r}
chessinfo.name <-rename(chess.info, c("Player.Name"="Player Name", "Player.State"="State", "Pre.Rating"="Pre-Rating", "Round.1"="Round 1", "Round.2"="Round 2", "Round.3"="Round 3", "Round.4"="Round 4", "Round.5"="Round 5", "Round.5.1"="Round 6", "Round.7"="Round 7", "Total.Points"="Ttl. Pts.", "Ave..Op..Rating"="Ave.Opp.Rating"))
head(chessinfo.name)
```

*After renaming, I would like to create a subset that contains just the columns we need for the CSV. We don't need to include each opponent's pre-rating.*

```{r}
chess.tournament <- chessinfo.name[, c(1,2,3,11,12)]
head(chess.tournament)
```

*Finally, to create the .CSV file.*
```{r}
write.csv(chess.tournament, file='chess_final.csv')
```



